Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-02-27 manuscript/manuscript.Rmd
Moved some simulation-related code to simulations.R

manuscript_dir <- file.path("C:/Users/alexz/Documents/GitHub/power.transform/manuscript")

@@

if(!file.exists(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))){

      shift_range <- 10^seq(from=0, to=6, by=0.1)

      for(d in shift_range){
        for(distribution in c("normal", "right-skewed", "left-skewed")){
          if(distribution == "normal"){
            x <- x_normal

          } else if(distribution == "left-skewed"){
            x <- x_left_skewed
          }

          for(method in c("box_cox", "yeo_johnson")){
            for(shift in c(FALSE, TRUE)){
              for(robust in c(FALSE)){
                for(estimation_method in setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())){

                  # Skip if robust, but not shifted.
                  if(robust && !shift) next

                  if(!shift && !robust){
                    version <- "original"

                  } else if(!robust){
                    version <- "shift-sensitive"

                  } else {
                    version <- "robust shift-sensitive"
                  }

                  yield(list(
                    "x" = x + d,
                    "d" = d,
                    "distribution" = distribution,
                    "method" = method,
                    "shift" = shift,
                    "robust" = robust,
                    "version" = version,
                    "estimation_method" = estimation_method
                  ))
                }
              }
            }
          }
        }
      }
    }
  )

  .compute_lambda <- function(parameter_set){
    # Create transformer object.
    if(parameter_set$version == "original"){
      transformer <- suppressWarnings(
      power.transform::find_transformation_parameters(
        x = parameter_set$x,
        method = parameter_set$method,
        robust = parameter_set$robust,
        shift = parameter_set$shift,
        estimation_method = parameter_set$estimation_method,
        lambda = NULL,
        optimiser_control = list("xtol_rel"=1e-5)))

    } else {
      transformer <- suppressWarnings(
      power.transform::find_transformation_parameters(
        x = parameter_set$x,
        method = parameter_set$method,
        robust = parameter_set$robust,
        shift = parameter_set$shift,
        estimation_method = parameter_set$estimation_method))
    }

    return(
      data.table::data.table(
        "distribution" = parameter_set$distribution,
        "method" = parameter_set$method,
        "estimation_method" = parameter_set$estimation_method,
        "version" = parameter_set$version,
        "d" = log10(parameter_set$d),
        "lambda" = transformer@lambda,
        "x_0" = transformer@shift
      ))
  }

  #### Computations --------------------------------------------------------------

  # data <- lapply(
  #   X = experiments,
  #   FUN = .compute_lambda)

  # Start cluster
  cl <- parallel::makeCluster(16L)

  # Save to file.
  saveRDS(data, file.path(manuscript_dir, "shifted_distributions_plot.RDS"))

} else {
  data <- readRDS(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))
}

# Only evaluate the MLE method here.
data <- data[estimation_method == "mle"]

# Update distribution, method and version to factors.
data$distribution <- factor(
  x = data$distribution,
  levels = c("normal", "right-skewed", "left-skewed"))
data$method <- factor(
  x = data$method,
  levels = c("box_cox", "yeo_johnson"),
  labels = c("Box-Cox", "Yeo-Johnson"))
data$version <- factor(
  x = data$version,
  levels = c("original", "shift-sensitive"))

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}

# Generate experiments.
experiment_args <- coro::generator(
  function(
    manuscript_dir,
    transformation_methods = NULL,
    estimation_methods = NULL){

    if(is.null(transformation_methods)){
      transformation_methods <- c("box_cox", "yeo_johnson")
    }

    if(is.null(estimation_methods)){
      estimation_methods <- setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())
    }

    for(estimation_method in estimation_methods){

      for(transformation_method in transformation_methods){
        # Non-robust transformation.
        fun_args <- list(
          "method" = transformation_method,
          "estimation_method" = estimation_method,
          "robust" = FALSE,
          "shift" = TRUE)

        file_name <- file.path(
          manuscript_dir,
          paste0(
            "robustness_comparison_",
            transformation_method, "_",
            estimation_method, "_", "non_robust.RDS"))

        if(!file.exists(file_name)){
          # Only yield arguments if the file does not exist.
          yield(list(
            "name" = "non-robust",
            "method" = transformation_method,
            "estimation_method" = estimation_method,
            "file_name" = file_name,
            "fun_args" = fun_args,
            "opt_args" = opt_args))
        } else {
          yield(file_name)
        }
      }

      for(robustness_source in c("empirical_probability", "transformed", "residual")){

        if(robustness_source == "empirical_probability"){
          def_opt_limits$k1[2] <- 1.0
          def_opt_limits$k2[2] <- 1.0

            fun_args <- list(
              "method" = transformation_method,
              "estimation_method" = estimation_method,
              "robust" = TRUE,
              "shift" = TRUE,
              "weighting_function" = paste0(robustness_source, "_", robustness_weighting_function),
              "backup_use_default" = FALSE)

            file_name <- file.path(
              manuscript_dir,
              paste0(
                "robustness_comparison_",
                transformation_method, "_",
                estimation_method, "_",
                robustness_source, "_",
                robustness_weighting_function, ".RDS"))

            if(!file.exists(file_name)){
              # Only yield arguments if the file does not exist.
              yield(list(
                "name" = paste0(robustness_source, "-", robustness_weighting_function),
                "method" = transformation_method,
                "estimation_method" = estimation_method,
                "file_name" = file_name,
                "fun_args" = fun_args,
                "opt_args" = list("initial" = opt_init, "limits" = opt_limits)))

            } else {
              yield(file_name)
            }
          }
        }
      }
    }
  }
)

# Helper function for computing the target lambda value.
.compute_target_lambda <- function(
    x,
    transformation_methods = NULL,
    estimation_methods = NULL){

  if(is.null(transformation_methods)){
    transformation_methods <- c("box_cox", "yeo_johnson")
  }

  if(is.null(estimation_methods)){
    estimation_methods <- setdiff(power.transform:::..estimators_all(), power.transform:::..estimators_raymaekers_robust())
  }

  lambda_list <- list()
  ii <- 1L
  for(transformation_method in transformation_methods){
    for(estimation_method in estimation_methods){
      # Create transformer object.
      transformer <- suppressWarnings(
        power.transform::find_transformation_parameters(
          x = x,
          method = transformation_method,
          estimation_method = estimation_method,
          robust = FALSE,
          shift = TRUE))

      lambda_list[[ii]] <- data.table::data.table(
        "method" = transformation_method,
        "estimation_method" = estimation_method,
        "lambda" = transformer@lambda)

      ii <- ii + 1L
    }
  }

  return(data.table::rbindlist(lambda_list))
}

# Helper function for population distributions with outliers.
.populate_outliers <- function(
    x,
    n,
    alpha,
    beta,
    target_lambda,
    outlier_fraction){

  # Set parameters.
  parameters <- list(
    "n" = n,
    "alpha" = alpha,
    "beta" = beta,
    "target_lambda" = target_lambda)

  # Add outliers.
  x <- lapply(
    outlier_fraction,
    function(k, x, parameters){
      # Set parameters.
      parameters$k <- k

  return(x)
}

# Helper function for computing transformer parameters under optimisation
# constraints.
.compute_robust_lambda <- function(
    x,
    fun_args,
    opt_args){

  # Custom parser.
  ..outlier_parser <- function(
    x,
    fun_args,
    opt_args){
    # Create transformer.
    transformer <- suppressWarnings(do.call(
      power.transform::find_transformation_parameters,
      args=c(
        list("x" = x$x),
        list("weighting_function_parameters" = opt_args),
        fun_args)))

    target_lambda <- x$parameters$target_lambda[method == fun_args$method & estimation_method == fun_args$estimation_method]$lambda

    parameter_data <- c(
      opt_args,
      list(
        "method" = fun_args$method,
        "estimation_method" = fun_args$estimation_method,
        "weight_method" = fun_args$weight_method,
        "lambda" = transformer@lambda,
        "target_lambda" = target_lambda,
        "shift" = transformer@shift)
    )

  parameter_data <- lapply(
    x,
    ..outlier_parser,
    fun_args = fun_args,
    opt_args = opt_args)

# Outer optimisation function used to optimise the weighting function parameters
# over a dataset.
.optimisation_outer <- function(
    experiment,
    data,
    verbose = FALSE){

  results_list <- list(
    "name" = experiment$name,
    "method" = experiment$method,
    "estimation_method" = experiment$estimation_method)

    h <- nloptr::sbplx(
      x0 = x0,
      fn = .optimisation_inner,
      lower = lower,
      upper = upper,
      control=list("xtol_rel"=1e-2),
      x = data,
      fun_args = experiment$fun_args,
      verbose = verbose)

    results_list <- c(results_list, list("k1" = h$par[1]))
    if(length(h$par) > 1) results_list <- c(results_list, list("k2" = h$par[2]))
    results_list <- c(results_list, list("value" = h$value))

  } else {
    value <- .optimisation_inner(
      opt_param=list(),
      x = data,
      fun_args=experiment$fun_args)

    results_list <- c(results_list, list("value" = value))
  }

  # Save to file.
  saveRDS(results_list, file=experiment$file_name)

  return(invisible(TRUE))
}

# Inner optimisation function that sets the loss.
.optimisation_inner <- function(
    opt_param,
    x,
    fun_args,
    verbose = FALSE){

  parameter_data <- .compute_robust_lambda(
    x = x,
    fun_args = fun_args,
    opt_args = opt_args)

  lambda_error <- sapply(
    parameter_data,
    function(x) (abs(x$lambda - x$target_lambda)))

  if(verbose){
    message <- paste0("target: ", mean(lambda_error))
    if(length(opt_args) > 0) message <- c(message, paste0("k1: ", opt_param[1]))
    if(length(opt_args) > 1) message <- c(message, paste0("k2: ", opt_param[2]))
    cat(paste0(message, collapse="; "))
    cat("\n")
  }

  return(mean(lambda_error))
}

# Select experiments to conduct by filtering out experiments that are file
# names.
experiments <- coro::collect(experiment_args(manuscript_dir = manuscript_dir))
experiments <- experiments[!sapply(experiments, is.character)]

if(length(experiments) > 0){
  set.seed(95)

  # Generate table of asymmetric generalised normal distribution parameters.
  n_distributions <- 3

  # Compute lambda values without weighting.
  target_lambda <- lapply(x, .compute_target_lambda)

  # Generate outliers in the data.
  x <- mapply(
    FUN = .populate_outliers,
    x = x,
    n = n,
    alpha = alpha,
    beta = beta,
    target_lambda = target_lambda,
    MoreArgs=list("outlier_fraction"=outlier_fraction),
    SIMPLIFY=FALSE,
    USE.NAMES=FALSE)

  lapply(experiments, .optimisation_outer, data=x, verbose=TRUE)

  parallel::clusterExport(
    cl=cl,
    varlist = c(".compute_robust_lambda", ".optimisation_inner")
  )

  parallel::parLapplyLB(
    cl = cl,
    X = experiments,
    fun = .optimisation_outer,
    data = x,
    chunk.size = 1L)

  # Stop cluster.
  parallel::stopCluster(cl)
}

# Collect all files
experiments <- coro::collect(experiment_args(manuscript_dir = manuscript_dir))
optimisation_data <- lapply(experiments, readRDS)
optimisation_data <- lapply(optimisation_data, data.table::as.data.table)
optimisation_data <- data.table::rbindlist(optimisation_data, fill=TRUE)
```

@@

if(!file.exists(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))){
  stop("Data are missing for creating shifted-distribution plots.")

} else {
  data <- readRDS(file.path(manuscript_dir, "shifted_distributions_plot.RDS"))
}

# Update distribution, method and version to factors.
data$distribution <- factor(
  x = data$distribution,
  levels = c("normal", "right-skewed", "left-skewed"))
data$method <- factor(
  x = data$method,
  levels = c("box_cox", "yeo_johnson"),
  labels = c("Box-Cox", "Yeo-Johnson"))
data$version <- factor(
  x = data$version,
  levels = c("original", "shift-sensitive"))
data$estimation_method <- factor(
  x = data$estimation_method,
  levels = c("mle", "anderson_darling", "cramer_von_mises", "jarque_bera", "dagostino"),
  labels = c("MLE", "Anderson-Darling", "Cramér-von Mises", "Jarque-Bera", "D'Agostino")
)

>>>>

manuscript_dir <- file.path("C:/Users/alexz/Documents/GitHub/power.transform/manuscript")
source(file.path(manuscript_dir, "simulations.R"))

@@

# Load data.
data <- .get_shifted_distribution_data(manuscript_dir=manuscript_dir)
data[estimation_method == "MLE"]

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
two_sided_function_parameters <- .get_optimised_weighting_function_parameters_two_sided(manuscript_dir = manuscript_dir)

```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
two_sided_transformation_parameters <- .get_transformation_parameters_two_sided(manuscript_dir = manuscript_dir)

@@

data <- .get_shifted_distribution_data(manuscript_dir = manuscript_dir)
